\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[most]{tcolorbox}
\usepackage{array}
\usepackage{latexsym}
\usepackage{alltt}
\usepackage{hyperref}
\usepackage{color}
\usepackage{float}
\usepackage{pdfpages}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{caption}
\usepackage{xparse}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{parskip}

\geometry
{
  a4paper,
  left=15mm,
  right=15mm,
  top=15mm,
  bottom=15mm,
}

% parskip
\setlength{\parindent}{16pt}

% multicolumns
\setlength{\columnsep}{1.6\columnsep}
\setlength{\columnseprule}{0.2pt}

% mybox
\newtcolorbox{mybox}[3][]
{
  colframe = #2!25,
  colback  = #2!10,
  coltitle = #2!20!black,  
  title    = {#3},
  #1,
}

% New environments that use mybox
\newcounter{example}
\newenvironment{example}[1]{\begin{mybox}{green}{\refstepcounter{example}\textbf{Example~\theexample #1}}}{\end{mybox}}

\newenvironment{examplebreak}[1]{\begin{mybox}[breakable]{green}{\refstepcounter{example}\textbf{Example~\theexample #1}}}{\end{mybox}}

\newcounter{definition}
\newenvironment{definition}[1]{\refstepcounter{definition}\begin{mybox}{blue}{\textbf{Definition~\thedefinition #1}}}{\end{mybox}}

\newcounter{theorem}
\newenvironment{theorem}[1]{\begin{mybox}{red}{\refstepcounter{theorem}\textbf{Theorem~\thetheorem #1}}}{\end{mybox}}

\newenvironment{formula}[1]{\begin{mybox}{cyan}{\textbf{#1}}}{\end{mybox}}

% Changing maketitle
\makeatletter         
\renewcommand\maketitle{
{\raggedright % Note the extra {
\begin{center}
{\Large \bfseries \@title}\\[2ex] 
{\large \@author \ - \@date}\\[2ex]
\end{center}}} % Note the extra }
\makeatother

% \onehalfspacing % adjust spacing

% macros
\newcommand{\prob}[1]{\textbf{\textit{P}}\{#1\}}
\NewDocumentCommand{\dsum}{%
    e{^_}
}{%
  {% 
    \displaystyle\sum
    \IfValueT{#1}{^{#1}}
    \IfValueT{#2}{_{#2}}
  }
}%

% maketitle variables
\title{CENG 222 - Chapter 8: Introduction to Statistics}
\author{Burak Metehan Tunçel}
\date{April 2022}

\begin{document}

\maketitle

\begin{multicols}{2}
  We learned analyzing problems and systems involving uncertainty, to find \textit{probabilities, expectations}, and \textit{other characteristics} for a variety of situations, and to produce forecasts that may lead to important decisions.

  What was given to us in all these problems? Ultimately, \textit{we needed to know the distribution and its parameters}, in order to compute probabilities or at least to estimate them by means of Monte Carlo. Often the distribution may not be given, and we learned how to fit the suitable model, say, Binomial, Exponential, or Poisson, given the type of variables we deal with. In any case, parameters of the fitted distribution had to be reported to us explicitly, or they had to follow directly from the problem.

  This, however, is rarely the case in practice. Only sometimes the situation may be under our control, where, for example, produced items have predetermined specifications, and therefore, one knows parameters of their distribution.

  Much more often \textit{parameters are not known}. To apply the knowledge we learned, \textit{we need to collect data}. A properly collected sample of data can provide rather sufficient information about parameters of the observed system. We will learn how to use this sample
  \begin{itemize}
    \item to visualize data, understand the patterns, and make quick statements about the system's behavior;
    \item to characterize this behavior in simple terms and quantities;
    \item to estimate the distribution parameters;
    \item to assess reliability of our estimates;
    \item to test statements about parameters and the entire system;
    \item to understand relations among variables;
    \item to fit suitable models and use them to make forecasts
  \end{itemize}
\end{multicols}

\input{sections/section8.1.tex}

\end{document}
