\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[most]{tcolorbox}
\usepackage{array}
\usepackage{latexsym}
\usepackage{alltt}
\usepackage{hyperref}
\usepackage{color}
\usepackage{float}
\usepackage{pdfpages}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{caption}
\usepackage{xparse}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{pdflscape}


\geometry
{
  a4paper,
  left=5mm,
  right=5mm,
  top=5mm,
  bottom=5mm,
}

% mybox
\newtcolorbox{mybox}[3][]
{
  colframe = #2!25,
  colback  = #2!10,
  coltitle = #2!20!black,  
  title    = {#3},
  #1,
}

% New environments that use mybox
\newenvironment{example}[1]{\begin{mybox}{green}{\textbf{Example #1}}}{\end{mybox}}
\newenvironment{example_break}[1]{\begin{mybox}[breakable]{green}{\textbf{Example #1}}}{\end{mybox}}
\newenvironment{definition}[1]{\begin{mybox}{blue}{\textbf{Definition #1}}}{\end{mybox}}
\newenvironment{theorem}[1]{\begin{mybox}{red}{\textbf{Theorem #1}}}{\end{mybox}}
\newenvironment{formula}[1]{\begin{mybox}{red}{\textbf{#1}}}{\end{mybox}}

% Changing maketitle
\makeatletter         
\renewcommand\maketitle{
{\raggedright % Note the extra {
\begin{center}
{\Large \bfseries \@title}\\[2ex] 
{\large \@author \ - \@date}\\[2ex]
\end{center}}} % Note the extra }
\makeatother

% \onehalfspacing % adjust spacing

% macros
\newcommand{\prob}[1]{\textbf{\textit{P}}\{#1\}}
\NewDocumentCommand{\dsum}{%
    e{^_}
}{%
  {% 
    \displaystyle\sum
    \IfValueT{#1}{^{#1}}
    \IfValueT{#2}{_{#2}}
  }
}%

% maketitle variables
\title{Title}
\author{Burak Metehan Tunçel}
\date{\today}

\begin{document}

\section*{Student Information}

Name : Burak Metehan Tunçel\\
ID : 2468726\\

\section*{Answer 1}

\subsection*{a)}

The regular expression:
\begin{equation*}
  \left( \left( a \cup b \right)^* aa \left(a \cup b \right)^* bb \left( a \cup b \right)^* \right) \cup \left( \left( a \cup b \right)^* bb \left(a \cup b \right)^* aa \left( a \cup b \right)^* \right)
\end{equation*}
This is equivalent to:
\begin{equation*}
  (a \cup b)^* ((aa (a \cup b)^* bb) \cup (bb (a \cup b)^* aa)) (a \cup b)^*
\end{equation*}

\subsection*{b)}

\input{1b-NFA-constructing.tex}

\newpage
\subsection*{c)}

\input{1c_DFA_constructing.tex}

\subsection*{d)}

\input{1d-Tracing.tex}



\newpage
\section*{Answer 2}

\subsection*{a)}

\quad If $L_1$ is a regular language, it should satisfy pumping lemma. According to pumping lemma, if $L_1$ is regular, there should be an integer $p \geq 1$ such that $w = a^mb^n \in L_1$ with $|w| \geq p$ can be written as $w = xyz$ such that:
\begin{itemize}
  \item $|y| > 0$,
  \item $|xy| < p$, and
  \item $xy^iz \in L_1$ for each $i \geq 0$.
\end{itemize}

\noindent When we focus on $y$, there are 4 possible $y$, while $k, l \geq 1$:
\begin{enumerate}
  \item $y = e$
  \item $y = a^k$
  \item $y = b^l$
  \item $y = a^kb^l$
\end{enumerate}

\noindent We can notice that there is not $y$ such that satisfies the conditions of pumping lemma. 
\begin{enumerate}
  \item $y \neq e$, from pumping lemma itself. (Contradiction)
  \item $y \neq a^k$, $y$ cannot include \textit{some} $a$: There are some situations that is valid but not all situation is valid. If the $m = n + 1$, $m \ngtr n$ when $i = 0$ for all $y = a^k$, while $k \geq 1$. (Contradiction)
  \item $y \neq b^l$, $y$ cannot include \textit{some} $b$: $m \ngtr n$ when for some $i > 0$. (Contradiction)
  \item $y \neq a^kb^l$, $y$ cannot include both \textit{some} $a$ and \textit{some} $b$: New strings out of pattern are generated (Example: $aaabbbaabb$). (Contradiction)
\end{enumerate}

\noindent Since all possible $y$ cause contradictions (at some points), which means that there is no integer $p$ that $L_1$ is \textbf{not regular language} according to pumping lemma. According to \textit{Theorem 2.3.1} in the textbook, languages accepted by finite automata are closed under complementation. So, $\overline{L_1}$ is also \textbf{not regular} language which is $L_2$

So, $L_2$ is \textbf{not regular} language.

\subsection*{b)}

\subsubsection*{$L_4$ is not regular}

\quad $L_4$ defines a language whose strings include equal number of $a$'s and $b$'s where $a$'s are located before $b$'s. We can use pumping lemma to show that $L_4$ is not regular.

Assume $L_4$ is regular languages. Then there should be an integer $p \geq 1$ such that $w = a^nb^n \in L_4$ with $|w| \geq p$ can be written as $w = xyz$ such that:
\begin{itemize}
  \item $|y| > 0$,
  \item $|xy| < p$, and
  \item $xy^iz \in L_4$ for each $i \geq 0$.
\end{itemize}

If $y = a^i$ for some $i > 0$, $xz = a^{n-i}b^n \notin L_4$ for $y^0 = e$, contradicting theorem. In other words, there is not any $y$ that satisfies $xy^iz \in L_4$ for each $i \geq 0$. New strings, $xy^iz$, either are out of pattern or does not have same number of $a$'s and $b$'s.

So, $L_4$ is \textbf{not regular} language according to pumping lemma.

\subsubsection*{$L_5$ is regular}

\quad In $L_5$, it stated that $m, n \in \mathbb{N}$, which means $m, n \geq 0$. So, $L_5$ can be expressed as:
\begin{equation*}
  L_5 = a^*b^*
\end{equation*}
because of the definition of \href{https://en.wikipedia.org/wiki/Kleene_star}{Kleene star} (Kleene star basically means \textit{zero or more occurences}).

So, it means that $L_5$ has a regular expression. If there is certain regular expression for a language, that language is regular language (It is also stated book in page 50: \textit{Regular languages are all languages that can be described by regular expressions.}).

Therefore, $L_5$ is \textbf{regular}.

\subsubsection*{$L_6$ is regular}

\quad Again, since $L_6$ is represented by certain regular expression, $L_6$ is regular.

\subsubsection*{$L_4 \cup L_5 \cup L_6$ is regular}

\quad At that point, we know that:
\begin{itemize}
  \item $L_4$ is not regular
  \item $L_5$ is regular
  \item $L_6$ is regular
\end{itemize}

If all of them was regular, we can say that the union of them is also regular because the languages are closed under union. However, $L_4$ is not regular.

When we focus on $L_4 \cup L_5$, we can notice that $L_4$ is subset of $L_5$, $L_4 \subset L_5$. So their union should be $L_5$:
\begin{equation*}
  L_4 \cup L_5 = L_5
\end{equation*}
So
\begin{equation*}
  L_4 \cup L_5 \cup L_6 = L_5 \cup L_6
\end{equation*}

So, according to \textit{Theorem 2.3.1} in the textbook, languages accepted by finite automata are closed under union. So, $L_5 \cup L_6$ is also \textbf{regular}.

Finally, we can say that $L_4 \cup L_5 \cup L_6$ is \textbf{regular}.

\end{document}
